A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-18 shanghai.tex
Moving to a different computer. Still working on Part 2.

In Section~\ref{sec:GN} we introduce N\'eron models of tori and the Greenberg transform, which produce
group schemes over $\Fq$ from algebraic tori $T$ over $K$.
Section~\ref{sec:bdchar} then uses Theorem~\ref{thm:snake} to give a geometrization of characters of $T$
of bounded depth.
We give a reinterpretation of this result in Section~\ref{sec:quasichar}, using a limit process to simultaneously
geometrize all quasicharacters of $T$.  Sections~\ref{sec:bdchar} and \ref{sec:quasichar} also include
functoriality results that we hope to use in a future discussion of class field theory.
Finally, we close the paper in Section~\ref{sec:transfer} by discussing a transfer principle, which allows
a geometric transfer of characters between characteristic $p$ tori to characteristic $0$ tori.

@@

\part{Application to characters of algebraic tori over local fields}

Recall that $\Fq$ is a finite field.
Let $K$ be a non-archimedean local field with residue field $\Fq$.
We place no restriction on the characteristic of $K$. 
We write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$. 
%For any positive integer $d$, set $A_d\ceq \OK/\pK^d$.
Let $T$ be an algebraic torus over $K$.
We write $\Hom_{< d}(T(K),\EEx)$ for the group of characters of $T(K)$ depth less than $d$ 
and $\Hom_{<\infty}(T(K),\EEx)$ for the group of characters of $T(K)$ of arbitrary finite depth;
one might call these quasicharacters of $T(K)$, expanding on \cite{tate:67a}*{2.3}.

@@

  Next we show that this isomorphism is functorial.
The first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor,
while the second part shows that Trace of Frobenius is a natural transformation
$t: F_1 \to F_2$. When further combined with Lemma~\ref{prop:product},
we see that $F_1$ is an additive functor and $t: F_1 \to F_2$ is a
natural transformation between additive functors.
\end{proof}

@@

By Proposition~\ref{prop:pullback}, the surjective morphism $f_d : \GN{T}_{d+1}\to \GN{T}_d$ of group schemes induces a functor
\[
f_d^* : \QC(\GN{T}_d) \to \QC(\GN{T}_{d+1}).
\]

\begin{proposition}
The functor $f_d^* : \QC(\GN{T}_d) \to \QC(\GN{T}_{d+1})$ is full and faithful.
\end{proposition}

\begin{proof}
It suffices to find a morphism of schemes $i_d : \GN{T}_{d}\hookrightarrow \GN{T}_{d+1}$ such that $f_d\circ i_d = \id_{\GN{T}_d}$. Base change along $\OK/\pK^{d+1} \to \OK/\pK^d$ determines a closed immersion $\TT_{d+1} \hookrightarrow \TT_d$, though not a morphism of group schemes. Applying the Greenberg functor produces a closed immersion of schemes $i_d : \GN{T}_{d} \hookrightarrow \GN{T}_{d+1}$. Although $i_d$ is not a morphism of group schemes, it is a section of $f_d : \GN{T}_{d+1}\to \GN{T}_d$.
\end{proof}

Let $\QC(\GN{T})$ be the category of ind-objects built from the categories $\QC(\GN{T}_d)$. 
An object $\qcs{F}\in \QC(\GN{T})$ is an equivalence class in the disjoint union of objects $\qcs{L}\in \GN{T}_d$, 
as $d$ ranges over the positive integers, 
under the relation that identifies $\qcs{L}\in \QC(\GN{T}_{d})$ with $f_d^*\qcs{L}\in \QC(\GN{T}_{d+1})$.
For each $\qcs{F}\in \QC(\GN{T})$ there is a unique least positive integer $d$ and a unique isomorphism class $\qcs{L}_{/\iso}\in \QCiso{\GN{T}_d}$ such that $\qcs{F} = [\qcs{L}]$; we refer to $d$ as the \emph{level} of $\qcs{F}$.
A map $\alpha : [\qcs{L}] \to [\qcs{L}']$ in $\QC(\GN{T})$ is

@@

\begin{definition}
The \emph{Greenberg transform of the N\'eron model of $T$} is the pro-group scheme over $\Fq$ given by
\begin{equation}\label{projlim}
  \cdef{\GN{T}} \ceq \varprojlim_{d\in \ZZ_{>0}} \GN{T}_{d}.
\end{equation}
\end{definition}

>>>>

In Section~\ref{sec:GN} we introduce N\'eron models of tori and the Greenberg transform, which produce
group schemes over $\Fq$ from algebraic tori $T$ over $K$.
Section~\ref{sec:bdchar} then uses Theorem~\ref{thm:snake} to give a geometrization of characters of $T$
of bounded depth.
We give a reinterpretation of this result in Section~\ref{sec:quasichar}, using a limit process to simultaneously
geometrize all quasicharacters of $T$.  Sections~\ref{sec:bdchar} and \ref{sec:quasichar} also include
functoriality results that we hope to use in a future discussion of class field theory.
\Clifton{There is another section at the end of the paper, now.}
Finally, we close the paper in Section~\ref{sec:transfer} by discussing a transfer principle, which allows
a geometric transfer of characters between characteristic $p$ tori to characteristic $0$ tori.

@@

\part{Application to quasicharacters of algebraic tori over local fields}

Recall that $\Fq$ is a finite field.
Let $K$ be a non-archimedean local field with residue field $\Fq$.
We place no restriction on the characteristic of $K$. 
We write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$. 
\Clifton{I think we should seriously consider replacing $d$ with $n$ in Part~2.}
%For any positive integer $d$, set $A_d\ceq \OK/\pK^d$.
Let $T$ be an algebraic torus over $K$.
We write $\Hom_{< d}(T(K),\EEx)$ for the group of characters of $T(K)$ depth less than $d$ 
and $\Hom_{<\infty}(T(K),\EEx)$ for the group of characters of $T(K)$ of arbitrary finite depth;
we call these quasicharacters of $T(K)$, following \cite{tate:67a}*{2.3}.

@@

  Next we show that this isomorphism is functorial.
  \Clifton{This proof isn't finished!}
The first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor,
while the second part shows that Trace of Frobenius is a natural transformation
$t: F_1 \to F_2$. When further combined with Lemma~\ref{prop:product},
we see that $F_1$ is an additive functor and $t: F_1 \to F_2$ is a
natural transformation between additive functors.
\end{proof}

@@

For any $d\leq d'$, define $f_{d\leq d'} : \GN{T}_{d'} \to \GN{T}_{d}$ by $f_{d\leq d'}\ceq f_d \circ f_{d+1} \circ \cdots \circ f_{d'-1}$.
Let $f_{d\leq d'}^* : \QC(\GN{T}_d) \to \QC(\GN{T}_{d'})$ be the functor promised by Proposition~\ref{prop:pullback}.

\begin{proposition}
The functor $f_{d\leq d'}^* : \QC(\GN{T}_d) \to \QC(\GN{T}_{d'})$ is full and faithful.
\end{proposition}

\begin{proof}
It suffices to find a morphism of schemes $i_d : \GN{T}_{d}\hookrightarrow \GN{T}_{d+1}$ such that $f_d\circ i_d = \id_{\GN{T}_d}$. Base change along $\OK/\pK^{d+1} \to \OK/\pK^d$ determines a closed immersion of schemes $\TT_{d+1} \hookrightarrow \TT_d$ (not a morphism of group schemes). Applying the Greenberg functor produces a closed immersion of schemes $i_d : \GN{T}_{d} \hookrightarrow \GN{T}_{d+1}$. Although $i_d$ is not a morphism of group schemes, it is a section of $f_d : \GN{T}_{d+1}\to \GN{T}_d$.
\end{proof}

\begin{definition}
The \emph{Greenberg transform of the N\'eron model of $T$} is the pro-group scheme over $\Fq$ given by $\cdef{\GN{T}} \ceq \varprojlim_{d\in \ZZ_{>0}} \GN{T}_{d}$.
Let $\QC(\GN{T})$ be the following category. An object in $\QC(\GN{T})$ is an equivalence class of objects from $\QC(\GN{T}_d)$, as $d$ ranges over the positive integers, where $\qcs{L}_d \equiv \qcs{L}_{d'}$ if $\qcs{L}_{d'} = f_{d\leq d'}^* \qcs{L}_d$.

A map $\alpha : [\qcs{L}] \to [\qcs{L}']$ in $\QC(\GN{T})$ is
\end{definition}

For each $\qcs{F}\in \QC(\GN{T})$ there is a unique least positive integer $d$ and a unique isomorphism class $\qcs{L}_{/\text{iso}}\in \QCiso{\GN{T}_d}$ such that $\qcs{F} = [\qcs{L}]$; we refer to $d$ as the \emph{level} of $\qcs{F}$.

@@
